---
title: "Saison"
author: '- BB; KS -'
date: "July 13, 2017"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
```

```{r results="asis", echo = FALSE}
cat("
<style>
body {background-color: #FFFFFF;}
</style>
")
```

TECH SPECS

* [Inkbird Itc-308 Digital Temperature Controller](https://www.amazon.com/gp/product/B011296704/ref=oh_aui_detailpage_o03_s00?ie=UTF8&psc=1)
* [Zoo Med Repti Therm Under Tank Heater](https://www.amazon.com/gp/product/B0002AQCLO/ref=as_li_tl?ie=UTF8&camp=1789&creative=390957&creativeASIN=B0002AQCL4&linkCode=as2&tag=birdingninja-20&linkId=KD3LQQCUIAKX4GWM&th=1)
* A regular, shitty, no-frills refrigerator

FERMENTATION SCHEDULE

1. Brewed
2. Soured for two days using just reptile pad (two days was too sour)
3. Boiled hops and pitched yeast at 68
4. 3 days later started bumping temp up 1 degree/day until 76 (should have done immediately)
5. Held at 76 for a week 
6. Brought back down to 71 using 1 degree/day
7. Bottle