---
title: "Steam Beer
"
author: 'Authors: BB, KS'
date: 7/01/2018
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
```

```{r results="asis", echo = FALSE}
cat("
<style>
body {background-color: #FFFFFF;}
</style>
")
```

#Style Summary:

* OG: 1.048 - 1.054  
* FG: 1.011 - 1.014  
* IBUs: 30 - 45  
* SRM: 10 - 14  
* ABV: 4.5% - 5.5% 

More details can be found here: [BJCP Style Details](https://www.bjcp.org/docs/2015_Guidelines_Beer.pdf)  

A small excerpt from the source about this style:    
*"Overall Impression: A lightly fruity beer with firm, grainy maltiness, interesting toasty and caramel flavors, and showcasing rustic, traditional American hop characteristics."* - BJCP -

```{r Kitchen Constants, echo = FALSE}
evap_rate = 10 # Percent
mash_eff = 73.5 # Percent
trub_loss = 0.125 # Gal
water_shrinkage = 4 # Percent
equip_loss = 0.125 # Gal
fermentation_loss = 0.25 # 1.0325 # Gal...I know we "documented" this, but it's hard to believe
batch_size = 5 # Gal
```

#Kitchen Constants:

* Evaporation Rate = `r evap_rate`% 
* Mash Efficiency = `r mash_eff`%  
* Boil Kettle Dead Space = `r trub_loss` Gallons  
* Lauter Tun Dead Space = `r equip_loss` Gallons   
* Fermentation Loss = `r fermentation_loss` Gallons  

\* Small Note: *we will use math mainly from "Designing Great Beers" by Ray Daniels as it is intuitive and easy to follow for the modern day home brewer.*


```{r Recipe, echo= FALSE}
recipe <- data.frame(
  Recipe = "Steam Burr",
  Style = "California Common",
  Date = '7/1/2018',
  OG = 1.051,
  FG = 1.013 
)

kable(recipe)
```

## Fermentables

[Link to Potential Gravity From Grain Values](http://beersmith.com/Grains/Grains/GrainList.htm)

```{r Fermentables, echo= FALSE}
Ing1 <- data.frame(Grains = "German Pils Malt", 
                   Lbs = 8, 
                   Potential_SG = 1.037,
                   EZ_Water_Code = 6,
                   SRM = 2.0
                   )
Ing2 <- data.frame(Grains = "Crystal malt (40 °L)", 
                   Lbs = 1, 
                   Potential_SG = 1.034,
                   EZ_Water_Code = 10,
                   SRM = 40.0
                   )

Ing3 <- data.frame(Grains = "Victory", 
                   Lbs = 0.5, 
                   Potential_SG = 1.034,
                   EZ_Water_Code = 8,
                   SRM = 25.0
                   )


fermentables <- rbind(Ing1,Ing2,Ing3)
tot_Lbs <- sum(fermentables$Lbs)

fermentables <- mutate(fermentables, Percent = Lbs/tot_Lbs)

kable(fermentables)
```


Theoretical OG from this is:

```{r Gravity Equation, echo = TRUE, eval = FALSE}
# We still need to recreate the static data database
# For now we will just use raw values I looked up

Ingredient_Gravity = lbs_needed * Gravity_per_lb_malt * Mash Efficiency
```

```{r OG Calculation, echo = TRUE, eval = TRUE}
# Don't forget mash efficiency
fermentables <- fermentables %>% mutate(Ingredient_Gravity = Lbs*(Potential_SG-1)*1000*(mash_eff/100))

# Predicted OG
predicted_og <- 1+(sum(fermentables$Ingredient_Gravity)/batch_size/1000)
print(paste("Predicted OG:",predicted_og))

```

## Hops
```{r Hops, echo= FALSE}

hop1 <- data.frame(Hop = "Northern Brewer", Hop_Type = "pellet", Alpha_Acid_Percent = 0.08, Weight_Oz = 0.75, Boil_Time = 60)
hop2 <- data.frame(Hop = "Northern Brewer", Hop_Type = "pellet", Alpha_Acid_Percent = 0.08, Weight_Oz = 0.75, Boil_Time = 20)
hop3 <- data.frame(Hop = "Northern Brewer", Hop_Type = "pellet", Alpha_Acid_Percent = 0.08, Weight_Oz = 1.0, Boil_Time = 0)

hops <- rbind(hop1,hop2,hop3)

# values from Designing Great Beers pg. 80

Utilization_Table <- data.frame(
  boil_time = c(0,9,19,29,44,59,74,75),
  whole_hop = c(0,5,12,15,19,22,24,27)/100, #percent stored as decimal
  pellet_hop = c(0,6,15,19,24,27,30,34)/100
)

# calc utilization column
# ranges are {0, 1-9, 10-19, 20-29, 30-44, 45-59, 60-74, 75+}
findUtilization <- function(boilTime = 0, hopType = "pellet"){

  BT <- boilTime
  if(BT > 75){
  boil_time_bin <- 75
  }else{
  boil_time_bin <- filter(Utilization_Table, boil_time >= BT) %>% select(boil_time) %>% min()
  }  
  
  if(hopType == "pellet"){
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(pellet_hop)
  }else{
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(whole_hop)
  }
  
  return(as.numeric(Util))
  
}

# Vectorize findUtilization: https://stackoverflow.com/questions/21818181/applying-a-function-to-every-row-of-a-table-using-dplyr
# Apparently to define a new column that depends on 1 or more other columns sent into a custom function with values passed in a row by row fashion, then you need a vectorized function.

FindUtil <- Vectorize(findUtilization, vectorize.args = c("boilTime", "hopType"))
hops <- mutate(hops, Utilization = FindUtil(Boil_Time, Hop_Type))

# Calculate any change to IBU yeild constant Cgravity which represents correction
# for worts that have a gravity above 1.050.

Cgravity <- if_else(predicted_og < 1.050, 1, 1 + (predicted_og - 1.050)/0.2)

hops <- hops %>% mutate(IBU = (Weight_Oz)*(Utilization)*(Alpha_Acid_Percent)*7489 / ((batch_size)*(Cgravity)) )

kable(hops)

print(paste("Total IBU:", sum(hops$IBU)))

```

## Mash

```{r Theoretical Values, echo=TRUE, eval=FALSE}
MV = Mash Volume [Gal]   ;EL = Equip Loss [Gal]          ;WS = Wort Shrinkage [%]
BS = Batch Size [Gal]    ;MT = Mash Thickness [Qts/Lbs]  ;GA = Grain Absorption [Qts/Lbs]
GB = Grain Bill [Lbs]    ;ST = Strike Temp [F]           ;ER = Evap Rate [%/hr]
BT = Boil Time [Min]     ;GT = Grain Temp [F]            ;VFR = Volume of First Runnings [Gal]
TL = Trub Loss [Gal]     ;TMT = Target Mash Temp [F]     ;AW = Addition Water in Mash [Gal]
SV = Sparge Volume [Gal] ;TWN = Total Water Needed [Gal] ;PBV = Pre Boil Volume [Gal]
IT = Infusion Temp [F]   ;MOV = Mash Out Volume [Gal]    ;SVPMO = Sparge Volume Post Mash Out [Gal]
MD = Mash Duration [Min] ;FL = Fermentation Loss [Gal]
```

```{r Equations for Mash Temps and Volumes, echo=TRUE, eval=FALSE}
# First we need to take batch size (BS) and inflate it by trub loss and fermentation loss
# Since we know the water will shrink after it cools, we then add on extra water to compensate (WS).
# Finally we use the Evaporatoin Rate to compensate for water lost due to boil.
# BT is in minutes so we divide by 60 to convert to hours.

# This gives us the Pre Boil Volume
PBV = ( (BS+(TL+FL))/(1-WS/100) ) / ( 1-(ER/100*(BT/60)) )

# Mash volume is simpley your mash thickness desired scaled by your grain amount.
MV = GB*(MT/4) # Convert to [Gal] for MT

# Technically the volume from your first runnings is used the accurately determine grain absorption.
# It also helps you find your equipment losses.
# If you directly measure your VFR you can skip this theoretical calculation and plug it into the next 
VFR = MV + AW - (GB*GA) - EL # Theoretical EQ...we could just measure

# In the end our pre boil volume minus our first runnings is the water we need to add to hit our batch size
SV = PBV - VFR

# Using a thermodynamic equation we can solve for the temperature that the mash volume needs to be
# in order to hit the the target mash temperature
IT = (0.2/MT)*(TMT-GT) + TMT

# If you so desire you can pull a specific volume from the SV and bring to a boil.
# Using this volume you can bring the mash sitting at TMT to mash out or 170 degrees fahrenheit.
# The equation to find this volume is below.
MOV = ( (170 - TMT)*(0.2*GB + MV) )/(212 - 170)

# Sparge Volume Post Mash Out
# May or may not be necessary; This is what is left of SV if we mash out with MOV
# We could just not pour in all of MOV and just sparge slowly with all of SV
# as long as it is hot enough to reach mash out.
# If not, we can heat with stove top to reach mash out.
SVPMO = SV - MOV 

# Also for convenience this is the Total Water Needed to achieve the batch size
TWN = MV + SV

```

Our recipe came out as follows:

```{r Mash, echo= FALSE, eval=TRUE}
# Initialize certain constants
BS <- batch_size
WS <- water_shrinkage
TL <- trub_loss
FL <- fermentation_loss
EL <- equip_loss
BT <- max(hops$Boil_Time) # Or hard code a longer mash time if that is the case
AW <- 0 # [Gal] Add if needed
ER <- evap_rate
MT <- 1.25 # Adjust if desired
GB <- sum(fermentables$Lbs)
GA <- 0.6 # https://byo.com/bock/item/410-calculating-water-usage-advanced-brewing uses 0.5
          # but to be conservative we will assume 0.6

GT <- 70
TMT <- 148
MD <- 60



PBV = ( (BS+(TL+FL))/(1-WS/100) ) / ( 1-(ER/100*(BT/60)) )

MV = GB*(MT/4) # Convert to [Gal] for MT

VFR = MV + AW - (GB*GA/4) - EL # Theoretical EQ...we could just measure

SV = PBV - VFR

IT = (0.2/MT)*(TMT-GT) + TMT

MOV = ( (170 - TMT)*(0.2*GB + MV) )/(212 - 170)

SVPMO = SV - MOV # May or may not be necessary; This is what is left of SV if we mash out with MOV
                 # We could just not pour in all of MOV and just sparge slowly with all of SV
                 # as long as it is hot enough to reach mash out.
                 # Even if not we can heat with stove top.

TWN = MV + SV

mash <- data.frame(
  Mash_Thickness = MT,
  Sacc_Rest_Temp = TMT,
  Mash_Duration = MD,
  Init_Grain_Temp = GT,
  Infusion_Temp = IT,
  Mash_Volume_Gal = MV,
  Sparge_Vol = SV,
  Mash_Out_Vol = MOV
)


kable(mash)
```

## Water Chemistry

Formulas and constants from [here](http://www.ezwatercalculator.com/)  
Bostons Water Supply Chemistry [link](https://www.brewersfriend.com/homebrew/water-profiles/)

```{r Water}

init_Ca <- 4
init_Mg <- 1
init_Na <- 32
init_Cl <- 23
init_SO4 <- 6
init_HCO3_CaCO3 <- 40

lactic_acid_percent <- 88 # %
lactic_acid_ml <- 0
acidulated_malt_acid_percent <- 2
acidulated_malt_oz <- 0

ph_down_gypsum_CaSO4 <- 5
ph_down_Cal_Chl_CaCl2 <- 5
ph_down_epsom_salt_MgSO4 <- 0
ph_up_slaked_lime_CaOH2 <- 0
ph_up_baking_soda_NaHCO3 <- 0
ph_up_chalk_CaCO3 <- 0

# Post mash additons
mash_Ca <- init_Ca + 
          (ph_up_chalk_CaCO3*105.89 +
           ph_down_gypsum_CaSO4*60 +
           ph_down_Cal_Chl_CaCl2*72 + 
           ph_up_slaked_lime_CaOH2*143)/ifelse(MV == 0, 1, MV)

mash_Mg <- init_Mg + 
          (ph_down_epsom_salt_MgSO4*24.6)/ifelse(MV == 0, 1, MV)

mash_Na <- init_Na + 
          (ph_up_baking_soda_NaHCO3*72.3)/ifelse(MV == 0, 1, MV)

mash_Cl <- init_Cl + 
          (ph_down_Cal_Chl_CaCl2*127.47)/ifelse(MV == 0, 1, MV)

mash_SO4 <- init_SO4 + 
          (ph_down_gypsum_CaSO4*147.4 +
           ph_down_epsom_salt_MgSO4*103)/ifelse(MV == 0, 1, MV)

# For the distilled water ph look up table
# 1 - Select Grains
# 2 - 2-Row
# 3 - 6-Row
# 4 - Maris Otter
# 5 - Munich
# 6 - Pilsner
# 7 - Wheat
# 8 - Vienna
# 9 - Other
# 10 - Crystal Malt
# 11 - Roasted/Toasted

dist_water_ph_look_up_table <- data.frame(EZ_Water_Code = 0:11,
                                          Distilled_Water_pH = c(0,
                                                                 0,
                                                                 5.7,
                                                                 5.79,
                                                                 5.77,
                                                                 5.43,
                                                                 5.75,
                                                                 6.04,
                                                                 5.56,
                                                                 5.70,
                                                                 NA,
                                                                 4.71))


fermentables <- left_join(x = fermentables, y = dist_water_ph_look_up_table, by = "EZ_Water_Code")

# Calculates the distilled water ph from crystal malts which are code 10
fermentables <- mutate(fermentables, Distilled_Water_pH = ifelse(EZ_Water_Code == 10, 5.22-0.00504*SRM,Distilled_Water_pH))

# Calculating Effective Alkalinity
if(MV == 0){
  
  effective_alkalinity <- 0
    
}else{

# There is this radio button for Bicarbonate versus Alkalinity
# We can't see how it works in the sheet.

ratio = 50/61
#ratio = 1

effective_alkalinity <- (init_HCO3_CaCO3*ratio+
  (
    (ph_up_chalk_CaCO3*130)+
    (ph_up_baking_soda_NaHCO3*157)-
    (176.1*lactic_acid_percent/100*lactic_acid_ml*2)-
    (4160.4*acidulated_malt_acid_percent/100*acidulated_malt_oz)*2.5+
    (ph_up_slaked_lime_CaOH2)*357
  )/(ifelse(MV == 0,1,MV)))

}
  
# Calculating Residual Alkalinity

if(MV == 0){
  residual_alkalinity <- 0
}else{
  residual_alkalinity <- effective_alkalinity - ((mash_Ca/1.4)+(mash_Mg/1.7))
            
}

# Estimated Room-Temp Mash pH
if(MV == 0){
  estimated_room_temp_mash_ph <- 0
}else{
  estimated_room_temp_mash_ph <- fermentables %>%
                                 mutate(Lbs_x_distWaterPH = Lbs*Distilled_Water_pH) %>%
                                 select(Lbs_x_distWaterPH) %>%
                                 sum()/ifelse(GB == 0, 1, GB) %>%
                                 as.numeric() + (0.1085*MV/ifelse(GB == 0, 1, GB)+0.013)*residual_alkalinity/50
}

# Choose to hard code or not.
# actual_ph <- 5.5
actual_ph <- estimated_room_temp_mash_ph

chemistry <- data.frame(
  recipe_id = 1,
  Init_Ca = init_Ca,
  Init_Mg = init_Mg,
  Init_Na = init_Na,
  Init_Cl = init_Cl,
  Init_SO4 = init_SO4,
  Init_HCO3_CaCO3 = init_HCO3_CaCO3,
  Actual_pH = actual_ph,
  Effective_Alkalinity = effective_alkalinity,
  Residual_Alkalinity = residual_alkalinity,
  pH_DOWN_Gypsum_CaSO4 = ph_down_gypsum_CaSO4,
  pH_DOWN_Cal_Chl_CaCl2 = ph_down_Cal_Chl_CaCl2,
  pH_DOWN_Epsom_Salt_MgSO4 = ph_down_epsom_salt_MgSO4,
  pH_UP_Slaked_Lime_CaOH2 = ph_up_slaked_lime_CaOH2,
  pH_UP_Baking_Soda_NaHCO3 = ph_up_baking_soda_NaHCO3,
  pH_UP_Chalk_CaCO3 = ph_up_chalk_CaCO3)

print(paste("Effective Alkalinity:", effective_alkalinity))
print(paste("Residual Alkalinity:", residual_alkalinity))
print(paste("Estimated Room-Temp Mash pH:", estimated_room_temp_mash_ph))
print(paste("Desired Room-Temp Mash pH: 5.4 - 5.6"))

kable(chemistry)

```

 
## Yeast
```{r Yeast, echo= FALSE}

# Using style OG
# degree_plato <- (recipe$OG-1)*1000/4
# Using current predicted OG
degree_plato <- (predicted_og-1)*1000/4

liter_batch_size <- batch_size*3.785 # Liters in a gallon

cell_density <- (1e9*degree_plato*liter_batch_size)

# Assumed that you want the right cell density and made it happen
pitched_cells <- cell_density

# IF otherwise, specify

# pitched_cells <- 150

yeast <- data.frame(
  Yeast_Strain = "WYEAST STRAIN: 2112 (California Lager)",
  Attenuation = 69,
  Initial_Cell_Count_Billions = cell_density/1e9,
  Pitched_Cells_Billions = pitched_cells/1e9,
  Liters_For_Starter = 0,
  Time_For_Starter = 0
)


predicted_fg <- predicted_og-(yeast$Attenuation/100)*(predicted_og-1)
predicted_abv <- (1.05/0.79)*((predicted_og-predicted_fg)/predicted_fg)*100

print(paste("Predicted ABV:", predicted_abv))

kable(yeast)
```


```{r Code to solve for starter volume needed, eval=FALSE, echo=TRUE}
# Code to solve for starter volume needed

I have some code that will bootleg this calculation and or some rules of thumb for starters...

```


## Fermentation

```{r Fermentation, echo= FALSE}

time1 <- data.frame(Temp = 60, Days = 21)
time2 <- data.frame(Temp = 45, Days = 7)


fermentation <- rbind(time1,time2)

kable(fermentation)
```












