---
title: "American IPA"
author: 'Authors: BB, KS'
date: 10/28/2018
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
```

```{r results="asis", echo = FALSE}
cat("
<style>
body {background-color: #FFFFFF;}
</style>
")
```

#Style Summary:

* OG:  1.056 - 1.070
* FG:  1.008 - 1.014
* ABV: 5.5 - 7.5
* IBU: 40 - 70
* SRM: 6 - 14

More details can be found here: [BJCP Style Details](https://beerrecipes.org/BJCP-2015-Style/65/21a-american-ipa.html)  

A small excerpt from the source about this style:    
*"Overall Impression: A decidedly hoppy and bitter, moderately strong American pale ale, showcasing modern American or New World hop varieties. The balance is hop-forward, with a clean fermentation profile, dryish finish, and clean, supporting malt allowing a creative range of hop character to shine through."* - BJCP -

```{r Kitchen Constants, echo = FALSE}
evap_rate = 15 # Percent
mash_eff = 73.5 # Percent
mash_tun_dead_space = 0 # Gal Mash Tun Dead Space
kettle_dead_space = 0.5 # Gal Kettle Deal Space
lauter_tun_dead_space = 0.5 # Gal Lauter Tun Dead Space
fermentation_tank_loss = 0.25 # Gal Fermentation Tank Loss
water_shrinkage = 4 # Percent
batch_size = 5 # Gal
```

#Kitchen Constants:

* Evaporation Rate = `r evap_rate`% 
* Mash Efficiency = `r mash_eff`%  
* Mash Tun Dead Space = `r mash_tun_dead_space` Gallons  
* Boil Kettle Dead Space = `r kettle_dead_space` Gallons  
* Lauter Tun Dead Space = `r lauter_tun_dead_space` Gallons   
* Fermentation Loss = `r fermentation_tank_loss` Gallons  

\* Small Note: *we will use math mainly from "Designing Great Beers" by Ray Daniels as it is intuitive and easy to follow for the modern day home brewer.*


```{r Recipe, echo= FALSE}
recipe <- data.frame(
  Recipe = "Juicy American IPA",
  Style = "American IPA",
  Date = '10_28_2018',
  OG = 1.063,
  FG = 1.011 
)

kable(recipe)
```

## Fermentables

[Link to Potential Gravity From Grain Values](http://beersmith.com/Grains/Grains/GrainList.htm)

```{r Fermentables, echo= FALSE}
Ing1 <- data.frame(Grains = "Pale Malt (2 Row) US", 
                   Lbs = 12.31, 
                   Potential_SG = 1.036,
                   EZ_Water_Code = 2,
                   SRM = 2.0
                   )


fermentables <- rbind(Ing1)
tot_Lbs <- sum(fermentables$Lbs)

fermentables <- mutate(fermentables, Percent = Lbs/tot_Lbs)

kable(fermentables)
```


Theoretical OG from this is:

```{r Gravity Equation, echo = TRUE, eval = FALSE}
# We still need to recreate the static data database
# For now we will just use raw values I looked up

Ingredient_Gravity = lbs_needed * Gravity_per_lb_malt * Mash Efficiency
```

```{r OG Calculation, echo = TRUE, eval = TRUE}
# Don't forget mash efficiency
fermentables <- fermentables %>% mutate(Ingredient_Gravity = Lbs*(Potential_SG-1)*1000*(mash_eff/100))

# Predicted OG
predicted_og <- 1+(sum(fermentables$Ingredient_Gravity)/batch_size/1000)
print(paste("Predicted OG:",predicted_og))

```

## Hops
So in the hop profile I want to avoid the resinous part of the hops addition or contribution to flavor. I think it is harsh. In general I want citrusy and fruity hops, but we will boil them to get the bitterness to come through. 

IBU is supposed to be in the range of 40-70. Here are some beers to model ours after:  
* Sierra Nevada Pale Ale: 38 IBU  
* Heady Topper: 75 IBU  
* Dogfish Head 90 min IPA: 90 IBU  

Citra boasts 60-70% of its hop oil volume in mercene which is known for its highly flavorful citrus and fruit flavors and or aromas. We will dry hop post fermentation so the yeast do not interfere with its flavor. Mosaic is also pretty flavorful (48-55% Mercene) and we will use it to bitter the beer. I believe that the isomerization of the hop oils when you boil leaves you with only bitterness flavor. Instead I intend to boil the hops for only 30 minutes. The achieve the correct IBU level, I will just use more hops. I believe that by not being conservative with the hop amount, I will achieve plenty of bitterness, while keeps the integrity of the hop flavors. After reading a bit about boiling hops, I’ve come the the conclusion that most flavor and or aroma is lost after boiling and it would never really matter what hop you used for boiling. Since I’m not trying to make the hoppiest beer in the world and flavor matters, keeping the characteristics of these hops intact is paramount. 

```{r Hops, echo= FALSE}

hop1 <- data.frame(Hop = "Mozaic", Hop_Type = "pellet", Alpha_Acid_Percent = 12.25/100, Weight_Oz = 2, Boil_Time = 30)
hop2 <- data.frame(Hop = "Citra", Hop_Type = "pellet", Alpha_Acid_Percent = 12.5/100, Weight_Oz = 0.75, Boil_Time = 5)
hop3 <- data.frame(Hop = "Citra", Hop_Type = "pellet", Alpha_Acid_Percent = 12.5/100, Weight_Oz = 0.75, Boil_Time = 0)

hops <- rbind(hop1,hop2,hop3)

# values from Designing Great Beers pg. 80

Utilization_Table <- data.frame(
  boil_time = c(0,9,19,29,44,59,74,75),
  whole_hop = c(0,5,12,15,19,22,24,27)/100, #percent stored as decimal
  pellet_hop = c(0,6,15,19,24,27,30,34)/100
)

# calc utilization column
# ranges are {0, 1-9, 10-19, 20-29, 30-44, 45-59, 60-74, 75+}
findUtilization <- function(boilTime = 0, hopType = "pellet"){

  BT <- boilTime
  if(BT > 75){
  boil_time_bin <- 75
  }else{
  boil_time_bin <- filter(Utilization_Table, boil_time >= BT) %>% select(boil_time) %>% min()
  }  
  
  if(hopType == "pellet"){
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(pellet_hop)
  }else{
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(whole_hop)
  }
  
  return(as.numeric(Util))
  
}

# Vectorize findUtilization: https://stackoverflow.com/questions/21818181/applying-a-function-to-every-row-of-a-table-using-dplyr
# Apparently to define a new column that depends on 1 or more other columns sent into a custom function with values passed in a row by row fashion, then you need a vectorized function.

FindUtil <- Vectorize(findUtilization, vectorize.args = c("boilTime", "hopType"))
hops <- mutate(hops, Utilization = FindUtil(Boil_Time, Hop_Type))

# Calculate any change to IBU yeild constant Cgravity which represents correction
# for worts that have a gravity above 1.050.

Cgravity <- if_else(predicted_og < 1.050, 1, 1 + (predicted_og - 1.050)/0.2)

hops <- hops %>% mutate(IBU = (Weight_Oz)*(Utilization)*(Alpha_Acid_Percent)*7489 / ((batch_size)*(Cgravity)) )

kable(hops)

print(paste("Total IBU:", sum(hops$IBU)))

```

## Mash

```{r Theoretical Values, echo=TRUE, eval=FALSE}
BS = Batch Size [Gal]
MV = Mash Volume [Gal]; AW = Addition Water in Mash [Gal]
VFR = Volume of First Runnings [Gal]
SV = Sparge Volume [Gal]

TWN = Total Water Needed [Gal]
PBV = Pre Boil Volume [Gal]
MOV = Mash Out Volume [Gal]
SVPMO = Sparge Volume Post Mash Out [Gal]

MTDS = Mash Tun Dead Space [Gal]
LTDS = Lauter Tun Dead Space [Gal]
KDS = Kettle Dead Space [Gal]
FTL = Fermentation Tank Loss [Gal]

GB = Grain Bill [Lbs]
GA = Grain Absorption [Qts/Lbs]
WS = Wort Shrinkage [%]
ER = Evap Rate [%/hr]
MD = Mash Duration [Min]
MT = Mash Thickness [Qts/Lbs]
BT = Boil Time [Min]

ST = Strike Temp [F]
GT = Grain Temp [F]
TMT = Target Mash Temp [F]
IT = Infusion Temp [F]

```

Our recipe came out as follows:

```{r Mash, echo= FALSE, eval=TRUE}
# Initialize certain constants

MTDS = mash_tun_dead_space
LTDS = lauter_tun_dead_space
KDS = kettle_dead_space
FTL = fermentation_tank_loss
BS <- batch_size
WS <- water_shrinkage
BT <- max(hops$Boil_Time) # Or hard code a longer mash time if that is the case
AW <- 0 # [Gal] Add if needed
ER <- evap_rate
MT <- 1.25 # Adjust if desired
GB <- sum(fermentables$Lbs)
GA <- 0.532 # https://byo.com/bock/item/410-calculating-water-usage-advanced-brewing uses 0.5
          # but to be conservative we will assume 0.6

GT <- 70
TMT <- 149
MD <- 60

BT = 60

PBV = ( (BS+KDS+FTL)/(1-WS/100) ) / ( 1-((ER/100)*(BT/60)) )

MV = GB*(MT/4) # Convert to [Gal] for MT

VFR = MV + AW - (GB*GA/4) - (MTDS+LTDS) # Theoretical EQ...we could just measure

SV = PBV - VFR

mash_protein_rest_temp = 168

MOV = ((0.2*GB)+(MV*4))*(((mash_protein_rest_temp-TMT)/(210-mash_protein_rest_temp))/4)

SVPMO = SV - MOV

IT = (0.2/MT)*(TMT-GT)+TMT

SVPMO = SV - MOV # May or may not be necessary; This is what is left of SV if we mash out with MOV
                 # We could just not pour in all of MOV and just sparge slowly with all of SV
                 # as long as it is hot enough to reach mash out.
                 # Even if not we can heat with stove top.

TWN = MV + SV

mash <- data.frame(
  Mash_Thickness = MT,
  Sacc_Rest_Temp = TMT,
  Mash_Duration = MD,
  Init_Grain_Temp = GT,
  Infusion_Temp = IT,
  Mash_Volume_Gal = MV,
  Sparge_Vol = SVPMO,
  Mash_Out_Vol = MOV
)


kable(mash)
```

## Water Chemistry

Formulas and constants from [here](http://www.ezwatercalculator.com/)  
Bostons Water Supply Chemistry [link](https://www.brewersfriend.com/homebrew/water-profiles/)

```{r Water}

init_Ca <- 4
init_Mg <- 1
init_Na <- 32
init_Cl <- 23
init_SO4 <- 6
init_HCO3_CaCO3 <- 40

lactic_acid_percent <- 88 # %
lactic_acid_ml <- 0
acidulated_malt_acid_percent <- 2
acidulated_malt_oz <- 0

ph_down_gypsum_CaSO4 <- 10
ph_down_Cal_Chl_CaCl2 <- 5
ph_down_epsom_salt_MgSO4 <- 5
ph_up_slaked_lime_CaOH2 <- 0
ph_up_baking_soda_NaHCO3 <- 0
ph_up_chalk_CaCO3 <- 0

# Post mash additons
mash_Ca <- init_Ca + 
          (ph_up_chalk_CaCO3*105.89 +
           ph_down_gypsum_CaSO4*60 +
           ph_down_Cal_Chl_CaCl2*72 + 
           ph_up_slaked_lime_CaOH2*143)/ifelse(MV == 0, 1, MV)

mash_Mg <- init_Mg + 
          (ph_down_epsom_salt_MgSO4*24.6)/ifelse(MV == 0, 1, MV)

mash_Na <- init_Na + 
          (ph_up_baking_soda_NaHCO3*72.3)/ifelse(MV == 0, 1, MV)

mash_Cl <- init_Cl + 
          (ph_down_Cal_Chl_CaCl2*127.47)/ifelse(MV == 0, 1, MV)

mash_SO4 <- init_SO4 + 
          (ph_down_gypsum_CaSO4*147.4 +
           ph_down_epsom_salt_MgSO4*103)/ifelse(MV == 0, 1, MV)

# For the distilled water ph look up table
# 1 - Select Grains
# 2 - 2-Row
# 3 - 6-Row
# 4 - Maris Otter
# 5 - Munich
# 6 - Pilsner
# 7 - Wheat
# 8 - Vienna
# 9 - Other
# 10 - Crystal Malt
# 11 - Roasted/Toasted

dist_water_ph_look_up_table <- data.frame(EZ_Water_Code = 0:11,
                                          Distilled_Water_pH = c(0,
                                                                 0,
                                                                 5.7,
                                                                 5.79,
                                                                 5.77,
                                                                 5.43,
                                                                 5.75,
                                                                 6.04,
                                                                 5.56,
                                                                 5.70,
                                                                 NA,
                                                                 4.71))


fermentables <- left_join(x = fermentables, y = dist_water_ph_look_up_table, by = "EZ_Water_Code")

# Calculates the distilled water ph from crystal malts which are code 10
fermentables <- mutate(fermentables, Distilled_Water_pH = ifelse(EZ_Water_Code == 10, 5.22-0.00504*SRM,Distilled_Water_pH))

# Calculating Effective Alkalinity
if(MV == 0){
  
  effective_alkalinity <- 0
    
}else{

# There is this radio button for Bicarbonate versus Alkalinity
# We can't see how it works in the sheet.

ratio = 50/61
#ratio = 1

effective_alkalinity <- (init_HCO3_CaCO3*ratio+
  (
    (ph_up_chalk_CaCO3*130)+
    (ph_up_baking_soda_NaHCO3*157)-
    (176.1*lactic_acid_percent/100*lactic_acid_ml*2)-
    (4160.4*acidulated_malt_acid_percent/100*acidulated_malt_oz)*2.5+
    (ph_up_slaked_lime_CaOH2)*357
  )/(ifelse(MV == 0,1,MV)))

}
  
# Calculating Residual Alkalinity

if(MV == 0){
  residual_alkalinity <- 0
}else{
  residual_alkalinity <- effective_alkalinity - ((mash_Ca/1.4)+(mash_Mg/1.7))
            
}

# Estimated Room-Temp Mash pH
if(MV == 0){
  estimated_room_temp_mash_ph <- 0
}else{
  estimated_room_temp_mash_ph <- fermentables %>%
                                 mutate(Lbs_x_distWaterPH = Lbs*Distilled_Water_pH) %>%
                                 select(Lbs_x_distWaterPH) %>%
                                 sum()/ifelse(GB == 0, 1, GB) %>%
                                 as.numeric() + (0.1085*MV/ifelse(GB == 0, 1, GB)+0.013)*residual_alkalinity/50
}

# Choose to hard code or not.
# actual_ph <- 5.5
actual_ph <- estimated_room_temp_mash_ph

chemistry <- data.frame(
  recipe_id = 1,
  Init_Ca = init_Ca,
  Init_Mg = init_Mg,
  Init_Na = init_Na,
  Init_Cl = init_Cl,
  Init_SO4 = init_SO4,
  Init_HCO3_CaCO3 = init_HCO3_CaCO3,
  Actual_pH = actual_ph,
  Effective_Alkalinity = effective_alkalinity,
  Residual_Alkalinity = residual_alkalinity,
  pH_DOWN_Gypsum_CaSO4 = ph_down_gypsum_CaSO4,
  pH_DOWN_Cal_Chl_CaCl2 = ph_down_Cal_Chl_CaCl2,
  pH_DOWN_Epsom_Salt_MgSO4 = ph_down_epsom_salt_MgSO4,
  pH_UP_Slaked_Lime_CaOH2 = ph_up_slaked_lime_CaOH2,
  pH_UP_Baking_Soda_NaHCO3 = ph_up_baking_soda_NaHCO3,
  pH_UP_Chalk_CaCO3 = ph_up_chalk_CaCO3)

print(paste("Effective Alkalinity:", effective_alkalinity))
print(paste("Residual Alkalinity:", residual_alkalinity))
print(paste("Estimated Room-Temp Mash pH:", estimated_room_temp_mash_ph))
print(paste("Desired Room-Temp Mash pH: 5.4 - 5.6"))

kable(chemistry)

```

 
## Yeast
```{r Yeast, echo= FALSE}

# Using style OG
# degree_plato <- (recipe$OG-1)*1000/4
# Using current predicted OG
degree_plato <- (predicted_og-1)*1000/4

liter_batch_size <- batch_size*3.785 # Liters in a gallon

cell_density <- (1e9*degree_plato*liter_batch_size)

# Assumed that you want the right cell density and made it happen
pitched_cells <- cell_density

# IF otherwise, specify

# pitched_cells <- 150

yeast <- data.frame(
  Yeast_Strain = "SafAle US-05",
  Attenuation = 72,
  Initial_Cell_Count_Billions = cell_density/1e9,
  Pitched_Cells_Billions = pitched_cells/1e9,
  Liters_For_Starter = 0,
  Time_For_Starter = 0
)


predicted_fg <- predicted_og-(yeast$Attenuation/100)*(predicted_og-1)
predicted_abv <- (1.05/0.79)*((predicted_og-predicted_fg)/predicted_fg)*100

print(paste("Predicted ABV:", predicted_abv))

kable(yeast)
```


```{r Code to solve for starter volume needed, eval=FALSE, echo=TRUE}
# Code to solve for starter volume needed

I have some code that will bootleg this calculation and or some rules of thumb for starters...

```


## Fermentation

We need to keep the hop character in the beer. This means we need to ferment somewhat cold for an ale or rather on the cold side of the yeast fermentation temperature range. Safale US-05 says 59 - 71.6 degrees fahrenheit. We will shoot for 59. As for any variation. The first week will be tightly controlled. Then we we will put a Kolsch in the fridge and the temp of the beer will probably average out to 70 degrees F for the remaining 3 weeks. 

```{r Fermentation, echo= FALSE}

time1 <- data.frame(Temp = 59, Days = 7)
time2 <- data.frame(Temp = 70, Days = 7)
time3 <- data.frame(Temp = 70, Days = 7)
time4 <- data.frame(Temp = 70, Days = 7)

fermentation <- rbind(time1,time2,time3,time4)

kable(fermentation)
```












