---
title: "Mexican Stout"
author: 'Authors: BB, KS'
date: "4/22/2018"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
```

```{r results="asis", echo = FALSE}
cat("
<style>
body {background-color: #FFFFFF;}
</style>
")
```

#Style Summary:

*If there’s one thing we can say without hesitation, it’s that imperial stout is expected to be big.* -BYO-

* OG: 1.100
* FG: 1.030
* IBUs: 60; BU:GU = 0.6 
* SRM: 500 - Black   
* ABV: 10.0%  

More details can be found here: [BJCP Style Details](https://www.bjcp.org/docs/2015_Guidelines_Beer.pdf)  

A small excerpt from the source about this style:    
*"Overall Impression: An intensely-flavored, big, dark ale with a wide range of flavor balances and regional interpretations. Roasty-burnt malt with deep dark or dried fruit flavors, and a warming, bittersweet finish. Despite the intense flavors, the components need to meld together to create a complex, harmonious beer, not a hot mess."* - BJCP -

*"Alcohol should be noticeable, contribute to flavor, and add a degree of smooth warming in the mouth and chest, but a hot character is a fault."* - BYO -

*"These also tend to be “full” beers, and unlike some other beers that might cheat by ramping up the carbonation to fill out the mouthfeel, imperial stouts come by it honestly by utilizing a significant percentage of caramel and chocolate malts in their grist. This tends to leave behind lots of unfermentable sugars, and the resulting texture should be solid and soft, but not syrupy."* -BYO-

*"Then we get into flavors, and suffice to say that they’re rarely subtle. Nearly all feature intense roast (sometimes bordering on burnt), especially in the form of the kind of high-cacao-percentage chocolate that tastes almost as much like the ground as it does a dessert. To this we usually add an array of dark pit fruits, dark caramel, rich bready melanoidins, and even piney or fruity hops. The best examples have a flavor profile that takes several seconds to fully play out."* -BYO-

*"The beer should have good body and taste like its aroma — rich, dark and chocolatey. The color should be pitch black and there should be a thick, brown head that forms. The best examples of imperial stout should have a balanced complexity, with all of the above attributes playing off each other without any overpowering the others."* -BYO-

Style Examples: [Examples](https://byo.com/article/the-king-of-stouts/)

```{r Kitchen Constants, echo = FALSE}
evap_rate = 10 # Percent
mash_eff = 60 # Percent
trub_loss = 0.125 # Gal
water_shrinkage = 4 # Percent
equip_loss = 0.125 # Gal
fermentation_loss = 0.25 # 1.0325 # Gal...I know we "documented" this, but it's hard to believe
batch_size = 5 # Gal
```

#Kitchen Constants:

* Evaporation Rate = `r evap_rate`% 
* Mash Efficiency = `r mash_eff`%  
* Boil Kettle Dead Space = `r trub_loss` Gallons  
* Lauter Tun Dead Space = `r equip_loss` Gallons   
* Fermentation Loss = `r fermentation_loss` Gallons  

\* Small Note: *we will use math mainly from "Designing Great Beers" by Ray Daniels as it is intuitive and easy to follow for the modern day home brewer.*


```{r Recipe, echo= FALSE}
options(warn=-1)
recipe <- data.frame(
  Recipe = "Mexican Stout",
  Style = "Imperial Stout",
  Date = Sys.Date(),
  OG = 1.100,
  FG = 1.030
)

kable(recipe)
```

## Fermentables

Notes from [learn.kegerator.com](https://learn.kegerator.com/russian-imperial-stout/):  
*Shoot for a grain to gallon ratio of about 3.5 pounds to every gallon. This means that the low end of a grain bill, for a 5 gallon batch, is going to weigh in at a hefty 17.5 pounds of grain and could go as high as 20 pounds.* 

*Your guiding standard should be complexity. Layer in complexity by using two, three, four, or five malts within each family.*

*Easily 20% of the brewing malts in your grain bill should come from these dark and roasted malts.* 

*"I like to start the grist with about 80–85 percent of a good bready base malt (Maris Otter is a good choice), and then fill in the gaps with a variety of light and dark crystals, some brown malt, and a lighter chocolate malt before capping it off with about 1–2 percent of an intense black malt (black patent or roasted barley)"* -BYO-

*"The most important part of brewing an imperial stout is the ingredients. I use only English Maris Otter for a base malt. We also layer our toasty and biscuity malts as well as our Cara malts. One is good but two, or three different malts in the same family are better — think complexity! "* -BYO-

*"Quality malts in general are best but the dark roasted grains are the most important to an imperial stout. The dark grains give the beer its chocolate and coffee flavors while the rest of the grains contribute to the final flavor. My opinion is that no less than seven malts should be used so that the complexity will not be compromised."* -BYO-

[Link to Potential Gravity From Grain Values](http://beersmith.com/Grains/Grains/GrainList.htm)
[More](http://www.brewunited.com/grain_database.php)

```{r Fermentables, echo= FALSE}
Ing1 <- data.frame(Grains = "Maris Otter 3.5 °L", 
                   Lbs = 20, 
                   Potential_SG = 1.038,
                   EZ_Water_Code = 2,
                   SRM = 4.0
                   )
Ing2 <- data.frame(Grains = "Cara Munich 71 °L", 
                   Lbs = 0.5, 
                   Potential_SG = 1.037,
                   EZ_Water_Code = 2,
                   SRM = 0.0
                   )
Ing3 <- data.frame(Grains = "Special B 120 °L", 
                   Lbs = 1, 
                   Potential_SG = 1.030,
                   EZ_Water_Code = 2,
                   SRM = 0.0
                   )
Ing4 <- data.frame(Grains = "Roasted Barley 500 °L", 
                   Lbs = 1.5, 
                   Potential_SG = 1.025,
                   EZ_Water_Code = 2,
                   SRM = 500.0
                   )
Ing5 <- data.frame(Grains = "Pale Chocolate Malt 200 °L", 
                   Lbs = 0.5, 
                   Potential_SG = 1.033,
                   EZ_Water_Code = 2,
                   SRM = 0.0
                   )
Ing6 <- data.frame(Grains = "Chocolate Malt 400 °L", 
                   Lbs = 0.5, 
                   Potential_SG = 1.028,
                   EZ_Water_Code = 2,
                   SRM = 0.0
                   )
fermentables <- rbind(Ing1,Ing2,Ing3,Ing4,Ing5,Ing6)
tot_Lbs <- sum(fermentables$Lbs)

fermentables <- mutate(fermentables, Percent = Lbs/tot_Lbs)

kable(fermentables)
```


Theoretical OG from this is:

```{r Gravity Equation, echo = TRUE, eval = FALSE}
# We still need to recreate the static data database
# For now we will just use raw values I looked up

Ingredient_Gravity = lbs_needed * Gravity_per_lb_malt * Mash Efficiency
```

```{r OG Calculation, echo = TRUE, eval = TRUE}
# Don't forget mash efficiency
fermentables <- fermentables %>% mutate(Ingredient_Gravity = Lbs*(Potential_SG-1)*1000*(mash_eff/100))

# Predicted OG
predicted_og <- 1+(sum(fermentables$Ingredient_Gravity)/batch_size/1000)
print(paste("Predicted OG:",predicted_og))

```

## Hops
From [learn.kegerator.com](https://learn.kegerator.com/russian-imperial-stout/):  
*The hops are not meant to standout though, and to this end it is important to use a hop with a relatively high alpha acid content, but one with little distinct flavor. Magnum, Northern Brewer, and East Kent Goldings are all good choices.*

*Put the bulk in as bittering, say 4/5th’s of your IBU number; at 60 minutes. Then the last 1/5th or so with only a minute left in the boil.*

*Basic no frills boil of 70 to 90 minutes. Boil over can be a concern, but again be mindful of the process and don’t try to hurry it.*

"*Bitterness varies, and you’ll want to throttle it based on how much of the darkest chocolate malt you use: More black malt should mean fewer IBUs, to avoid an overly-harsh flavor. Having said that, I prefer to go easier on the black malt and heavier on the hops, especially with some big mid-boil additions that will add bitterness and flavor. Play around with continental or old-school hops paired with fruity hops (Northern Brewer blended with Citra®, for example) to add bright fruit notes that will complement the plum and currant, while also adding some “mature” earthy and herbal notes. My IBU:OG ratio is about 1:1, which seems high, but keep in mind we’re talking about a 10% ABV beer, so the alcohol will be adding a lot of sweetness and we’ll be aging it for a while, which means some of the bitterness will drop out over time.*" -BYO-

*"The boil is pretty basic. It is up to the individual brewer, but in my experience 70 to 90 minutes is sufficient. As for boil-overs, they are always a concern — just pay attention and do not rush the boil and you should be fine."* -BYO-

*"I typically do two hop infusions in an imperial stout brewing session. One at about 60 minutes left in the boil and one at one minute left in the boil. The 60 minute addition should be between 2/3 and 4/5 of all the IBUs for the beer. The one minute addition should comprise the rest. In an imperial stout, I do not think that the hops are meant to stand out. They should play a supporting role to the dark malt flavors and aromas and keep the beer from being cloying."* -BYO-

*"Fifty to 60 IBUs is a lot of hops, but in this beer it is for balance."* -BYO-

*"BU:GU = 0.6-0.8"* -BYO-

Kyle\Ben
66% of hop IBU from bittering hop

```{r Hops, echo= FALSE}

hop1 <- data.frame(Hop = "Northern Brewer", Hop_Type = "pellet", Alpha_Acid_Percent = 0.085, Weight_Oz = 1.4, Boil_Time = 60)
hop2 <- data.frame(Hop = "Northern Brewer", Hop_Type = "pellet", Alpha_Acid_Percent = 0.085, Weight_Oz = 1.2, Boil_Time = 10)


hops <- rbind(hop1,hop2)

# values from Designing Great Beers pg. 80

Utilization_Table <- data.frame(
  boil_time = c(0,9,19,29,44,59,74,75),
  whole_hop = c(0,5,12,15,19,22,24,27)/100, #percent stored as decimal
  pellet_hop = c(0,6,15,19,24,27,30,34)/100
)

# calc utilization column
# ranges are {0, 1-9, 10-19, 20-29, 30-44, 45-59, 60-74, 75+}
findUtilization <- function(boilTime = 0, hopType = "pellet"){

  BT <- boilTime
  if(BT > 75){
  boil_time_bin <- 75
  }else{
  boil_time_bin <- filter(Utilization_Table, boil_time >= BT) %>% select(boil_time) %>% min()
  }  
  
  if(hopType == "pellet"){
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(pellet_hop)
  }else{
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(whole_hop)
  }
  
  return(as.numeric(Util))
  
}

# Vectorize findUtilization: https://stackoverflow.com/questions/21818181/applying-a-function-to-every-row-of-a-table-using-dplyr
# Apparently to define a new column that depends on 1 or more other columns sent into a custom function with values passed in a row by row fashion, then you need a vectorized function.

FindUtil <- Vectorize(findUtilization, vectorize.args = c("boilTime", "hopType"))
hops <- mutate(hops, Utilization = FindUtil(Boil_Time, Hop_Type))

# Calculate any change to IBU yeild constant Cgravity which represents correction
# for worts that have a gravity above 1.050.

Cgravity <- if_else(predicted_og < 1.050, 1, 1 + (predicted_og - 1.050)/0.2)

hops <- hops %>% mutate(IBU = (Weight_Oz)*(Utilization)*(Alpha_Acid_Percent)*7489 / ((batch_size)*(Cgravity)) )

kable(hops)

print(paste("Total IBU:", sum(hops$IBU)))

```

## Mash

From [learn.kegerator.com](https://learn.kegerator.com/russian-imperial-stout/):  
*If you can step mash, do it… if not shoot for a single temperature range between 152-158°F and mash for at least 90 minutes.*

*"First, mash at your normal temperature or a little below. I know we want a rich mouthfeel, but we should have plenty of residual sugars to get the job done. I mash this beer at my standard 152 °F (67 °C), and it more than gets the job done. You can go lower to promote attenuation, but don’t go higher. If you think it’s too thin (and you don’t detect hot alcohols, which might suggest a problem in fermentation), adjust the recipe and add more crystal malt or flaked grains to bulk up the beer. A note on crystal malts for body: If you go this route, you should be cautious that you’re not adding sweetness as well as body. Staying above the 40 °L threshold should keep added sweetness to a minimum, and you can also balance with a few more IBUs."* -BYO-

*"A mash of 90 minutes should be enough. Vorlauf for at least 20 minutes before collecting. The brewer should be trying to achieve the complexity in the mash. That is the reason for the 152–158 ºF (67–70 ºC) temperature range. It depends on how chewy you want it. The sign that it is working will be **when you finally start the runoff, the first runnings should look like used motor oil — pitch black and thick**."* -BYO-

```{r Theoretical Values, echo=FALSE, eval=FALSE}
MV = Mash Volume [Gal]   ;EL = Equip Loss [Gal]          ;WS = Wort Shrinkage [%]
BS = Batch Size [Gal]    ;MT = Mash Thickness [Qts/Lbs]  ;GA = Grain Absorption [Qts/Lbs]
GB = Grain Bill [Lbs]    ;ST = Strike Temp [F]           ;ER = Evap Rate [%/hr]
BT = Boil Time [Min]     ;GT = Grain Temp [F]            ;VFR = Volume of First Runnings [Gal]
TL = Trub Loss [Gal]     ;TMT = Target Mash Temp [F]     ;AW = Addition Water in Mash [Gal]
SV = Sparge Volume [Gal] ;TWN = Total Water Needed [Gal] ;PBV = Pre Boil Volume [Gal]
IT = Infusion Temp [F]   ;MOV = Mash Out Volume [Gal]    ;SVPMO = Sparge Volume Post Mash Out [Gal]
MD = Mash Duration [Min] ;FL = Fermentation Loss [Gal]
```

```{r Equations for Mash Temps and Volumes, echo=FALSE, eval=FALSE}
# First we need to take batch size (BS) and inflate it by trub loss and fermentation loss
# Since we know the water will shrink after it cools, we then add on extra water to compensate (WS).
# Finally we use the Evaporatoin Rate to compensate for water lost due to boil.
# BT is in minutes so we divide by 60 to convert to hours.

# This gives us the Pre Boil Volume
PBV = ( (BS+(TL+FL))/(1-WS/100) ) / ( 1-(ER/100*(BT/60)) )

# Mash volume is simpley your mash thickness desired scaled by your grain amount.
MV = GB*(MT/4) # Convert to [Gal] for MT

# Technically the volume from your first runnings is used the accurately determine grain absorption.
# It also helps you find your equipment losses.
# If you directly measure your VFR you can skip this theoretical calculation and plug it into the next 
VFR = MV + AW - (GB*GA) - EL # Theoretical EQ...we could just measure

# In the end our pre boil volume minus our first runnings is the water we need to add to hit our batch size
SV = PBV - VFR

# Using a thermodynamic equation we can solve for the temperature that the mash volume needs to be
# in order to hit the the target mash temperature
IT = (0.2/MT)*(TMT-GT) + TMT

# If you so desire you can pull a specific volume from the SV and bring to a boil.
# Using this volume you can bring the mash sitting at TMT to mash out or 170 degrees fahrenheit.
# The equation to find this volume is below.
MOV = ( (170 - TMT)*(0.2*GB + MV) )/(212 - 170)

# Sparge Volume Post Mash Out
# May or may not be necessary; This is what is left of SV if we mash out with MOV
# We could just not pour in all of MOV and just sparge slowly with all of SV
# as long as it is hot enough to reach mash out.
# If not, we can heat with stove top to reach mash out.
SVPMO = SV - MOV 

# Also for convenience this is the Total Water Needed to achieve the batch size
TWN = MV + SV

```

Our Mash Parameters:

```{r Mash, echo= FALSE, eval=TRUE}
# Initialize certain constants
BS <- batch_size
WS <- water_shrinkage
TL <- trub_loss
FL <- fermentation_loss
EL <- equip_loss
BT <- max(hops$Boil_Time) # Or hard code a longer mash time if that is the case
AW <- 0 # [Gal] Add if needed
ER <- evap_rate
MT <- 1.25 # Adjust if desired
GB <- sum(fermentables$Lbs)
GA <- 0.6 # https://byo.com/bock/item/410-calculating-water-usage-advanced-brewing uses 0.5
          # but to be conservative we will assume 0.6

GT <- 70
TMT <- 157
MD <- 60



PBV = ( (BS+(TL+FL))/(1-WS/100) ) / ( 1-(ER/100*(BT/60)) )

MV = GB*(MT/4) # Convert to [Gal] for MT

VFR = MV + AW - (GB*GA/4) - EL # Theoretical EQ...we could just measure

SV = PBV - VFR

IT = (0.2/MT)*(TMT-GT) + TMT

MOV = ( (170 - TMT)*(0.2*GB + MV) )/(212 - 170)

SVPMO = SV - MOV # May or may not be necessary; This is what is left of SV if we mash out with MOV
                 # We could just not pour in all of MOV and just sparge slowly with all of SV
                 # as long as it is hot enough to reach mash out.
                 # Even if not we can heat with stove top.

TWN = MV + SV

mash <- data.frame(
  Mash_Thickness = MT,
  Sacc_Rest_Temp = TMT,
  Mash_Duration = MD,
  Init_Grain_Temp = GT,
  Infusion_Temp = IT,
  Mash_Volume_Gal = MV,
  Sparge_Vol = SV,
  Mash_Out_Vol = MOV
)


kable(mash)
```

## Water Chemistry

Formulas and constants from [here](http://www.ezwatercalculator.com/)  
Bostons Water Supply Chemistry [link](https://www.brewersfriend.com/homebrew/water-profiles/)

```{r Water, echo=FALSE}

init_Ca <- 4
init_Mg <- 1
init_Na <- 32
init_Cl <- 23
init_SO4 <- 6
init_HCO3_CaCO3 <- 40

lactic_acid_percent <- 88 # %
lactic_acid_ml <- 0
acidulated_malt_acid_percent <- 2
acidulated_malt_oz <- 0

ph_down_gypsum_CaSO4 <- 0
ph_down_Cal_Chl_CaCl2 <- 0
ph_down_epsom_salt_MgSO4 <- 0
ph_up_slaked_lime_CaOH2 <- 0
ph_up_baking_soda_NaHCO3 <- 0
ph_up_chalk_CaCO3 <- 0

# Post mash additons
mash_Ca <- init_Ca + 
          (ph_up_chalk_CaCO3*105.89 +
           ph_down_gypsum_CaSO4*60 +
           ph_down_Cal_Chl_CaCl2*72 + 
           ph_up_slaked_lime_CaOH2*143)/ifelse(MV == 0, 1, MV)

mash_Mg <- init_Mg + 
          (ph_down_epsom_salt_MgSO4*24.6)/ifelse(MV == 0, 1, MV)

mash_Na <- init_Na + 
          (ph_up_baking_soda_NaHCO3*72.3)/ifelse(MV == 0, 1, MV)

mash_Cl <- init_Cl + 
          (ph_down_Cal_Chl_CaCl2*127.47)/ifelse(MV == 0, 1, MV)

mash_SO4 <- init_SO4 + 
          (ph_down_gypsum_CaSO4*147.4 +
           ph_down_epsom_salt_MgSO4*103)/ifelse(MV == 0, 1, MV)


dist_water_ph_look_up_table <- data.frame(EZ_Water_Code = 0:11,
                                          Distilled_Water_pH = c(0,
                                                                 0,
                                                                 5.7,
                                                                 5.79,
                                                                 5.77,
                                                                 5.43,
                                                                 5.75,
                                                                 6.04,
                                                                 5.56,
                                                                 5.70,
                                                                 NA,
                                                                 4.71))


fermentables <- left_join(x = fermentables, y = dist_water_ph_look_up_table, by = "EZ_Water_Code")

# Calculates the distilled water ph from crystal malts which are code 10
fermentables <- mutate(fermentables, Distilled_Water_pH = ifelse(EZ_Water_Code == 10, 5.22-0.00504*SRM,Distilled_Water_pH))

# Calculating Effective Alkalinity
if(MV == 0){
  
  effective_alkalinity <- 0
    
}else{

# There is this radio button for Bicarbonate versus Alkalinity
# We can't see how it works in the sheet.

ratio = 50/61
#ratio = 1

effective_alkalinity <- (init_HCO3_CaCO3*ratio+
  (
    (ph_up_chalk_CaCO3*130)+
    (ph_up_baking_soda_NaHCO3*157)-
    (176.1*lactic_acid_percent/100*lactic_acid_ml*2)-
    (4160.4*acidulated_malt_acid_percent/100*acidulated_malt_oz)*2.5+
    (ph_up_slaked_lime_CaOH2)*357
  )/(ifelse(MV == 0,1,MV)))

}
  
# Calculating Residual Alkalinity

if(MV == 0){
  residual_alkalinity <- 0
}else{
  residual_alkalinity <- effective_alkalinity - ((mash_Ca/1.4)+(mash_Mg/1.7))
            
}

# Estimated Room-Temp Mash pH
if(MV == 0){
  estimated_room_temp_mash_ph <- 0
}else{
  estimated_room_temp_mash_ph <- fermentables %>%
                                 mutate(Lbs_x_distWaterPH = Lbs*Distilled_Water_pH) %>%
                                 select(Lbs_x_distWaterPH) %>%
                                 sum()/ifelse(GB == 0, 1, GB) %>%
                                 as.numeric() + (0.1085*MV/ifelse(GB == 0, 1, GB)+0.013)*residual_alkalinity/50
}

# Choose to hard code or not.
# actual_ph <- 5.5
actual_ph <- estimated_room_temp_mash_ph

chemistry <- data.frame(
  recipe_id = 1,
  Init_Ca = init_Ca,
  Init_Mg = init_Mg,
  Init_Na = init_Na,
  Init_Cl = init_Cl,
  Init_SO4 = init_SO4,
  Init_HCO3_CaCO3 = init_HCO3_CaCO3,
  Actual_pH = actual_ph,
  Effective_Alkalinity = effective_alkalinity,
  Residual_Alkalinity = residual_alkalinity,
  pH_DOWN_Gypsum_CaSO4 = ph_down_gypsum_CaSO4,
  pH_DOWN_Cal_Chl_CaCl2 = ph_down_Cal_Chl_CaCl2,
  pH_DOWN_Epsom_Salt_MgSO4 = ph_down_epsom_salt_MgSO4,
  pH_UP_Slaked_Lime_CaOH2 = ph_up_slaked_lime_CaOH2,
  pH_UP_Baking_Soda_NaHCO3 = ph_up_baking_soda_NaHCO3,
  pH_UP_Chalk_CaCO3 = ph_up_chalk_CaCO3)

print(paste("Effective Alkalinity:", effective_alkalinity))
print(paste("Residual Alkalinity:", residual_alkalinity))
print(paste("Estimated Room-Temp Mash pH:", estimated_room_temp_mash_ph))
print(paste("Desired Room-Temp Mash pH: 5.4 - 5.6"))

kable(chemistry)

```

 
## Yeast
*Use twice the yeast you’d put in a normal brew. Oxygenate your wort well before casting the yeast; splash, stir, I’ve even used a blender on quantities. Danster Nottingham is a good choice in the dry yeast category and Wyeast 1028 is a great liquid yeast for this style.*

*Whatever the yeast, use two packets and make a good size starter. These beers can be monsters during the fermentation, creating a lot of excess heat, so it might be wise to store it in a temperature controlled refrigerator set at the low end of the yeast’s range.* -kegerator.com-

*"Finally, choose a yeast with a high alcohol tolerance and some clean ester character. I’d avoid the spicier Belgian strains entirely in favor of some of the heritage British strains or the Wyeast 1007 (German Ale) yeast. By all means, choose a yeast that will contribute some flavor, but not one that will compete with the rest of the recipe."* -BYO-

*"I can only recommend one yeast and that yeast is Wyeast 1028. I am familiar with this strain so I know what it will do. In my opinion it is the best yeast for the job. The best pitching temperature is between 66–72 ºF (19–22 ºC). Any ale strain should produce good results in this range. These beers are notoriously violent fermenters so hang on and enjoy a wild ride!"* -BYO-

*"I add a lager yeast at the end of fermentation for aging purposes. Dry-hopping is optional and it only takes a small amount as the beer has a lot of alcohol and body."* -BYO-

```{r Yeast, echo= FALSE}

# Using style OG
# degree_plato <- (recipe$OG-1)*1000/4
# Using current predicted OG
degree_plato <- (predicted_og-1)*1000/4

liter_batch_size <- batch_size*3.785 # Liters in a gallon

cell_density <- (1e9*degree_plato*liter_batch_size)

# Assumed that you want the right cell density and made it happen
pitched_cells <- cell_density

# IF otherwise, specify

# pitched_cells <- 150

yeast <- data.frame(
  Yeast_Strain = "Danstar Nottingham Ale Dry Yeast",
  Attenuation = 70,
  Initial_Cell_Count_Billions = cell_density/1e9,
  Pitched_Cells_Billions = pitched_cells/1e9,
  Liters_For_Starter = 0,
  Time_For_Starter = 0
)


predicted_fg <- predicted_og-(yeast$Attenuation/100)*(predicted_og-1)
predicted_abv <- (1.05/0.79)*((predicted_og-predicted_fg)/predicted_fg)*100

print(paste("Predicted ABV:", predicted_abv))

kable(yeast)
```


```{r Code to solve for starter volume needed, eval=FALSE, echo=TRUE}
# Code to solve for starter volume needed

I have some code that will bootleg this calculation and or some rules of thumb for starters...

```


## Fermentation

*"Second, ferment relatively cool and long, with plenty of yeast and oxygen. We want a clean fermentation. There will be more than enough alcohol presence without us roughing it up a bit, so make a starter (or increase the amount of purchased yeast) and get lots of oxygen (or air) into the chilled beer before pitching. Once you do that, start your fermentation at the low end of the ale range, about 64 °F (18 °F) and walk it up very slowly to help ensure full attenuation but limiting the production of hotter alcohols. I wouldn’t attempt this beer unless I had active temperature control or a highly temperature-stable basement (in winter). With the amount of simple sugars hanging around, the yeast will want to take off. We need to hold them in check, and then give them plenty of time to clean up when they’re done: After the completion of primary fermentation, leave it alone for a week (or two) to ensure that the yeast have given it their full attention."* -BYO-

*"Aging should be at cellar temperature or cooler."* -BYO-

```{r Fermentation, echo= FALSE}

time1 <- data.frame(Temp = 70, Days = 9)
time2 <- data.frame(Temp = 70, Days = 7)
time3 <- data.frame(Temp = 70, Days = 30)
time4 <- data.frame(Temp = 40, Days = 365*2) #Obviously we won't do 2 years lol

fermentation <- rbind(time1,time2,time3,time4)

kable(fermentation)
```












