---
title: "Peach Dunkelweissen"
author: 'Authors: BB, KS'
date: "October 26, 2017"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
```

```{r results="asis", echo = FALSE}
cat("
<style>
body {background-color: #FFFFFF;}
</style>
")
```

#Style Summary:

* OG: 1.065
* FG: 1.014  
* IBUs: 30
* SRM: 28
* ABV: 7%  

More details can be found here: [HomeBrewTalk](http://www.homebrewtalk.com/showthread.php?t=295301)  

```{r Kitchen Constants, echo = FALSE}
evap_rate = 10 # Percent
mash_eff = 73.5 # Percent
trub_loss = 0.125 # Gal
equip_loss = 0.125 # Gal
fermentation_loss = 0.25 # 1.0325 # Gal...I know we "documented" this, but it's hard to believe
water_shrinkage = 4 # Percent

batch_size = 7 # Gal
```

#Kitchen Constants:

* Evaporation Rate = 10% 
* Mash Efficiency = 73.5%  
* Boil Kettle Dead Space = 0.125 Gallons  
* Lauter Tun Dead Space = 0.125 Gallons   
* Fermentation Loss = 0.25 Gallons  

Small Note: *we will use math mainly from "Designing Great Beers" by Ray Daniels as it is intuitive and easy to follow for the modern day home brewer.*

#Our Recipe:

## Recipe Summary

```{r Recipe, echo= FALSE}
recipe <- data.frame(
  recipe_id = 1,
  Recipe = "Pecan Brew Bro",
  Style = "Dunkleweissen",
  Date = Sys.Date(),
  OG = 1.065,
  FG = 1.014 
)

kable(recipe[,2:6])
```

## Fermentables


```{r Fermentables, echo= FALSE}
fermentables <- data.frame(
  recipe_id = rep(1,6),
  Grains = c("White Wheat Malt",
             "Pilsner",
             "Cara Red", # Red Wheat substitute
             "Choc. Wheat",
             "Crystal 30", # Cara Wheat substitute
             "Honey Malt"),
  EZ_Water_Code = c(7,6,9,11,7,4),
  SRM = c(2,2,20,400,30,25),
  Lbs = 1.4*c(7, 2.5, 1, 0.5, 0.5, 0.75),
  Potential_SG = c(1.040,1.037,1.035,1.030,1.035,1.034)
  )


Lbs_Total <- sum(fermentables$Lbs)

fermentables["Percent"] <- 100*fermentables$Lbs/Lbs_Total

kable(fermentables[,2:6])
```

sources:  
- [grain constants](http://beersmith.com/Grains/Grains/GrainList.htm)  
- [honey malt constants](http://www.onebeer.net/grainchart.html)

```{r Gravity Equation, echo = FALSE, eval = FALSE}
# We still need to recreate the static data database
# For now we will just use raw values I looked up

Ingredient_Gravity = lbs_needed * Gravity_per_lb_malt * Mash Efficiency
```

```{r OG Calculation, echo = FALSE, eval = TRUE}
# Don't forget mash efficiency
fermentables <- fermentables %>% mutate(Ingredient_Gravity = Lbs*(Potential_SG-1)*1000*(mash_eff/100))

# Predicted OG
predicted_og <- 1+(sum(fermentables$Ingredient_Gravity)/(batch_size+fermentation_loss)/1000)
```

<strong>Theoretical OG from this is:</strong> `r predicted_og`

## Hops
```{r Hops, echo= FALSE}

# alpha acids from
# East Kent Goldings: https://ychhops.com/varieties/east-kent-golding
# Fuggle: https://ychhops.com/varieties/fuggle-1

hops <- data.frame(
  recipe_id = rep(1,2),
  Hop = c("Brewer's Gold","Brewer's Gold"),
  Hop_Type = c("pellet", "pellet"),
  Alpha_Acid_Percent = c(8.5/100,8.5/100), #percent stored as decimal
  Weight_Oz = c(.5,.5)*(9.5/8.5)*1.4,
  Boil_Time = c(60,15)
)

# values from Designing Great Beers pg. 80

Utilization_Table <- data.frame(
  boil_time = c(0,9,19,29,44,59,74,75),
  whole_hop = c(0,5,12,15,19,22,24,27)/100, #percent stored as decimal
  pellet_hop = c(0,6,15,19,24,27,30,34)/100
)

# calc utilization column
# ranges are {0, 1-9, 10-19, 20-29, 30-44, 45-59, 60-74, 75+}
findUtilization <- function(boilTime = 0, hopType = "pellet"){

  BT <- boilTime
  if(BT > 75){
  boil_time_bin <- 75
  }else{
  boil_time_bin <- filter(Utilization_Table, boil_time >= BT) %>% select(boil_time) %>% min()
  }  
  
  if(hopType == "pellet"){
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(pellet_hop)
  }else{
    Util <- filter(Utilization_Table, boil_time == boil_time_bin) %>% select(whole_hop)
  }
  
  return(as.numeric(Util))
  
}

# Vectorize findUtilization: https://stackoverflow.com/questions/21818181/applying-a-function-to-every-row-of-a-table-using-dplyr
# Apparently to define a new column that depends on 1 or more other columns sent into a custom function with values passed in a row by row fashion, then you need a vectorized function.

FindUtil <- Vectorize(findUtilization, vectorize.args = c("boilTime", "hopType"))
hops <- mutate(hops, Utilization = FindUtil(Boil_Time, Hop_Type))

# Calculate any change to IBU yeild constant Cgravity which represents correction
# for worts that have a gravity above 1.050.

Cgravity <- if_else(predicted_og < 1.050, 1, 1 + (predicted_og - 1.050)/0.2)

hops <- hops %>% mutate(IBU = (Weight_Oz)*(Utilization)*(Alpha_Acid_Percent)*7489 / ((batch_size)*(Cgravity)) )

kable(hops[,2:6])
```

<strong>Total IBU:</strong>  `r sum(hops$IBU)`

## Mash

8 oz Pecans --start with halves soaked overnight in salt water(to promote enzymes that improve flavor), slow roasted till dry, soaked overnight in vodka(to remove oils), slow roasted till dry, break up into quarters, leave over night, slow roast until barely starting to smoke, and then remove from oven, then add to grain bag and start wort. 45 min mashing.

- [Dunkelweissen Mash Details](https://www.northernbrewer.com/documentation/allgrain/AG-Dunkelweizen.pdf)

```{r Theoretical Values, echo=FALSE, eval=FALSE}
MV = Mash Volume [Gal]   ;EL = Equip Loss [Gal]          ;WS = Wort Shrinkage [%]
BS = Batch Size [Gal]    ;MT = Mash Thickness [Qts/Lbs]  ;GA = Grain Absorption [Qts/Lbs]
GB = Grain Bill [Lbs]    ;ST = Strike Temp [F]           ;ER = Evap Rate [%/hr]
BT = Boil Time [Min]     ;GT = Grain Temp [F]            ;VFR = Volume of First Runnings [Gal]
TL = Trub Loss [Gal]     ;TMT = Target Mash Temp [F]     ;AW = Addition Water in Mash [Gal]
SV = Sparge Volume [Gal] ;TWN = Total Water Needed [Gal] ;PBV = Pre Boil Volume [Gal]
IT = Infusion Temp [F]   ;MOV = Mash Out Volume [Gal]    ;SVPMO = Sparge Volume Post Mash Out [Gal]
MD = Mash Duration [Min] ;FL = Fermentation Loss [Gal]
```

```{r Equations for Mash Temps and Volumes, echo=FALSE, eval=FALSE}
# First we need to take batch size (BS) and inflate it by trub loss and fermentation loss
# Since we know the water will shrink after it cools, we then add on extra water to compensate (WS).
# Finally we use the Evaporatoin Rate to compensate for water lost due to boil.
# BT is in minutes so we divide by 60 to convert to hours.

# This gives us the Pre Boil Volume
PBV = ( (BS+(TL+FL))/(1-WS/100) ) / ( 1-(ER/100*(BT/60)) )

# Mash volume is simpley your mash thickness desired scaled by your grain amount.
MV = GB*(MT/4) # Convert to [Gal] for MT

# Technically the volume from your first runnings is used the accurately determine grain absorption.
# It also helps you find your equipment losses.
# If you directly measure your VFR you can skip this theoretical calculation and plug it into the next 
VFR = MV + AW - (GB*GA) - EL # Theoretical EQ...we could just measure

# In the end our pre boil volume minus our first runnings is the water we need to add to hit our batch size
SV = PBV - VFR

# Using a thermodynamic equation we can solve for the temperature that the mash volume needs to be
# in order to hit the the target mash temperature
IT = (0.2/MT)*(TMT-GT) + TMT

# If you so desire you can pull a specific volume from the SV and bring to a boil.
# Using this volume you can bring the mash sitting at TMT to mash out or 170 degrees fahrenheit.
# The equation to find this volume is below.
MOV = ( (170 - TMT)*(0.2*GB + MV) )/(212 - 170)

# Sparge Volume Post Mash Out
# May or may not be necessary; This is what is left of SV if we mash out with MOV
# We could just not pour in all of MOV and just sparge slowly with all of SV
# as long as it is hot enough to reach mash out.
# If not, we can heat with stove top to reach mash out.
SVPMO = SV - MOV 

# Also for convenience this is the Total Water Needed to achieve the batch size
TWN = MV + SV

```

Our recipe came out as follows:

```{r Mash, echo= FALSE, eval=TRUE}
# Initialize certain constants
BS <- batch_size
WS <- water_shrinkage
TL <- trub_loss
FL <- fermentation_loss
EL <- equip_loss
BT <- max(hops$Boil_Time) # Or hard code a longer mash time if that is the case
AW <- 0 # [Gal] Add if needed
ER <- evap_rate
MT <- 1.25 # Adjust if desired
GB <- sum(fermentables$Lbs)
GA <- 0.6 # https://byo.com/bock/item/410-calculating-water-usage-advanced-brewing uses 0.5
          # but to be conservative we will assume 0.6

# Graint Temp
GT <- 70
# Target Mash Temp
TMT <- 153
# Mash Duration
MD <- 60


# Pre-boil volume
PBV = ( (BS+(TL+FL))/(1-WS/100) ) / ( 1-(ER/100*(BT/60)) )
# Mash Volume
MV = GB*(MT/4) # Convert to [Gal] for MT
# Volume from runnings
VFR = MV + AW - (GB*GA/4) - EL # Theoretical EQ...we could just measure
# Sparge volume
SV = PBV - VFR
# Infusion temperature
IT = (0.2/MT)*(TMT-GT) + TMT
# Mash Out Volume
MOV = ( (170 - TMT)*((0.2*GB) + (MV*4)) )/(4*(212 - 170))  
# Sparge volume post mash out
SVPMO = SV - MOV # May or may not be necessary; This is what is left of SV if we mash out with MOV
                 # We could just not pour in all of MOV and just sparge slowly with all of SV
                 # as long as it is hot enough to reach mash out.
                 # Even if not we can heat with stove top.
# Total water needed
TWN = MV + SV

mash <- data.frame(
  recipe_id = 1,
  Mash_Thickness = MT,
  Sacc_Rest_Temp = TMT,
  Mash_Duration = MD,
  Init_Grain_Temp = GT,
  Infusion_Temp = IT,
  Mash_Volume_Gal = MV,
  Sparge_Vol = SV,
  Mash_Out_Vol = MOV
) # Remember Mash_Out_Vol is a subset of Sparge_Vol

# kable(mash[2:9])
```

<strong>Mash Thickness:</strong> `r MT`  
<strong>Sacc Rest Temp:</strong> `r TMT`  
<strong>Mash Duration:</strong> `r MD` min  
<strong>Initial Grain Temp:</strong> `r GT`  
<strong>Infusion Temp:</strong> `r IT`  
<strong>Mash Volume (Gal):</strong> `r MV`  
<strong>Sparge Volume (Gal):</strong> `r SV`  
<strong>Mash Out Volume (Gal):</strong> `r MOV` 


## Water Chemistry

Formulas and constants from [here](http://www.ezwatercalculator.com/)  
Bostons Water Supply Chemistry [link](https://www.brewersfriend.com/homebrew/water-profiles/)

```{r Water, echo=FALSE}

init_Ca <- 4
init_Mg <- 1
init_Na <- 32
init_Cl <- 23
init_SO4 <- 6
init_HCO3_CaCO3 <- 40

percent_mash_vol_distilled <- 0 # %
percent_sparge_vol_distilled <- 0 # %
lactic_acid_percent <- 88 # %
lactic_acid_ml <- 0
acidulated_malt_acid_percent <- 2
acidulated_malt_oz <- 0

# Mash Additions
ph_down_gypsum_CaSO4 <- 2; adjust_sparge_gypsum_CaSO4 <- TRUE
ph_down_Cal_Chl_CaCl2 <- 10; adjust_sparge_Cal_Chl_CaCl2 <- TRUE
ph_down_epsom_salt_MgSO4 <- 8; adjust_sparge_epsom_salt_MgSO4 <- TRUE
ph_up_slaked_lime_CaOH2 <- 2; adjust_sparge_slaked_lime_CaOH2 <- TRUE
ph_up_baking_soda_NaHCO3 <- 0; adjust_sparge_baking_soda_NaHCO3 <- TRUE
ph_up_chalk_CaCO3 <- 0; adjust_sparge_chalk_CaCO3 <- TRUE

# Sparge Additions (auto calculated)
s_ph_down_gypsum_CaSO4 <- ifelse(adjust_sparge_gypsum_CaSO4, ph_down_gypsum_CaSO4*SV/MV, 0)
s_ph_down_Cal_Chl_CaCl2 <- ifelse(adjust_sparge_Cal_Chl_CaCl2, ph_down_Cal_Chl_CaCl2*SV/MV, 0)
s_ph_down_epsom_salt_MgSO4 <- ifelse(adjust_sparge_epsom_salt_MgSO4, ph_down_epsom_salt_MgSO4*SV/MV, 0)
s_ph_up_slaked_lime_CaOH2 <- ifelse(adjust_sparge_slaked_lime_CaOH2, ph_up_slaked_lime_CaOH2*SV/MV, 0)
s_ph_up_baking_soda_NaHCO3 <- ifelse(adjust_sparge_baking_soda_NaHCO3, ph_up_baking_soda_NaHCO3*SV/MV, 0)
s_ph_up_chalk_CaCO3 <- ifelse(adjust_sparge_chalk_CaCO3, ph_up_chalk_CaCO3*SV/MV, 0)


# Post mash additions
mash_Ca <- init_Ca + 
          (ph_up_chalk_CaCO3*105.89 +
           ph_down_gypsum_CaSO4*60 +
           ph_down_Cal_Chl_CaCl2*72 + 
           ph_up_slaked_lime_CaOH2*143)/ifelse(MV == 0, 1, MV)

mash_Mg <- init_Mg + 
          (ph_down_epsom_salt_MgSO4*24.6)/ifelse(MV == 0, 1, MV)

mash_Na <- init_Na + 
          (ph_up_baking_soda_NaHCO3*72.3)/ifelse(MV == 0, 1, MV)

mash_Cl <- init_Cl + 
          (ph_down_Cal_Chl_CaCl2*127.47)/ifelse(MV == 0, 1, MV)

mash_SO4 <- init_SO4 + 
          (ph_down_gypsum_CaSO4*147.4 +
           ph_down_epsom_salt_MgSO4*103)/ifelse(MV == 0, 1, MV)

# Post mash and sparge additions
mash_sparge_Ca <- ifelse(SV==0,
                         mash_Ca,
                         (1-
                           ((
                             (percent_mash_vol_distilled*MV) +                                      
                             (percent_sparge_vol_distilled*SV)
                            )/(MV+SV)
                           )
                          )*init_Ca + 
                             ((ph_up_chalk_CaCO3+s_ph_up_chalk_CaCO3)*105.89 +         
                             (ph_down_gypsum_CaSO4+s_ph_down_gypsum_CaSO4)*60 + 
                             (ph_down_Cal_Chl_CaCl2+s_ph_down_Cal_Chl_CaCl2)*72 + 
                             (ph_up_slaked_lime_CaOH2+s_ph_up_slaked_lime_CaOH2)*143
                             )/(MV+SV)
                         )

mash_sparge_Mg <- ifelse(SV==0,
                         mash_Mg,
                         (1-
                           ((
                             (percent_mash_vol_distilled*MV) +                                      
                             (percent_sparge_vol_distilled*SV)
                            )/(MV+SV)
                           )
                          )*init_Mg + 
                             ((ph_down_epsom_salt_MgSO4+s_ph_down_epsom_salt_MgSO4)*24.6/(MV+SV)
                         )
                         )

mash_sparge_Na <- ifelse(SV==0,
                         mash_Na,
                         (1-
                           ((
                             (percent_mash_vol_distilled*MV) +                                      
                             (percent_sparge_vol_distilled*SV)
                            )/(MV+SV)
                           )
                          )*init_Na + 
                             ((ph_up_baking_soda_NaHCO3+s_ph_up_baking_soda_NaHCO3)*72.3/(MV+SV)
                         
                         )
                         )

mash_sparge_Cl <- ifelse(SV==0,
                         mash_Cl,
                         (1-
                           ((
                             (percent_mash_vol_distilled*MV) +                                      
                             (percent_sparge_vol_distilled*SV)
                            )/(MV+SV)
                           )
                          )*init_Cl + 
                             ((ph_down_Cal_Chl_CaCl2+s_ph_down_Cal_Chl_CaCl2)*127.47/(MV+SV)
                         
                         )
                         )

mash_sparge_SO4 <- ifelse(SV==0,
                         mash_SO4,
                         (1-
                           ((
                             (percent_mash_vol_distilled*MV) +                                      
                             (percent_sparge_vol_distilled*SV)
                            )/(MV+SV)
                           )
                          )*init_SO4 + 
                             ((ph_down_gypsum_CaSO4+s_ph_down_gypsum_CaSO4)*147.4 +         
                             (ph_down_epsom_salt_MgSO4+s_ph_down_epsom_salt_MgSO4)*103)/(MV+SV)
                         )


dist_water_ph_look_up_table <- data.frame(EZ_Water_Code = 0:11,
                                          Distilled_Water_pH = c(0,
                                                                 0,
                                                                 5.7,
                                                                 5.79,
                                                                 5.77,
                                                                 5.43,
                                                                 5.75,
                                                                 6.04,
                                                                 5.56,
                                                                 5.70,
                                                                 NA,
                                                                 4.71))


fermentables <- left_join(x = fermentables, y = dist_water_ph_look_up_table, by = "EZ_Water_Code")

# Calculates the distilled water ph from crystal malts which are code 10
fermentables <- mutate(fermentables, Distilled_Water_pH = ifelse(EZ_Water_Code == 10, 5.22-0.00504*SRM,Distilled_Water_pH))

# Calculating Effective Alkalinity
if(MV == 0){
  
  effective_alkalinity <- 0
    
}else{

# There is this radio button for Bicarbonate versus Alkalinity
# We can't see how it works in the sheet.

ratio = 50/61
#ratio = 1

effective_alkalinity <- (init_HCO3_CaCO3*ratio+
  (
    (ph_up_chalk_CaCO3*130)+
    (ph_up_baking_soda_NaHCO3*157)-
    (176.1*lactic_acid_percent/100*lactic_acid_ml*2)-
    (4160.4*acidulated_malt_acid_percent/100*acidulated_malt_oz)*2.5+
    (ph_up_slaked_lime_CaOH2)*357
  )/(ifelse(MV == 0,1,MV)))

}
  
# Calculating Residual Alkalinity

if(MV == 0){
  residual_alkalinity <- 0
}else{
  residual_alkalinity <- effective_alkalinity - ((mash_Ca/1.4)+(mash_Mg/1.7))
            
}

# Estimated Room-Temp Mash pH
if(MV == 0){
  estimated_room_temp_mash_ph <- 0
}else{
  estimated_room_temp_mash_ph <- fermentables %>%
                                 mutate(Lbs_x_distWaterPH = Lbs*Distilled_Water_pH) %>%
                                 select(Lbs_x_distWaterPH) %>%
                                 sum()/ifelse(GB == 0, 1, GB) %>%
                                 as.numeric() + (0.1085*MV/ifelse(GB == 0, 1, GB)+0.013)*residual_alkalinity/50
}

# Choose to hard code or not.
# actual_ph <- 5.5
actual_ph <- estimated_room_temp_mash_ph

chemistry <- data.frame(
  recipe_id = 1,
  Init_Ca = init_Ca,
  Init_Mg = init_Mg,
  Init_Na = init_Na,
  Init_Cl = init_Cl,
  Init_SO4 = init_SO4,
  Init_HCO3_CaCO3 = init_HCO3_CaCO3,
  Actual_pH = actual_ph,
  Effective_Alkalinity = effective_alkalinity,
  Residual_Alkalinity = residual_alkalinity,
  pH_DOWN_Gypsum_CaSO4 = ph_down_gypsum_CaSO4,
  pH_DOWN_Cal_Chl_CaCl2 = ph_down_Cal_Chl_CaCl2,
  pH_DOWN_Epsom_Salt_MgSO4 = ph_down_epsom_salt_MgSO4,
  pH_UP_Slaked_Lime_CaOH2 = ph_up_slaked_lime_CaOH2,
  pH_UP_Baking_Soda_NaHCO3 = ph_up_baking_soda_NaHCO3,
  pH_UP_Chalk_CaCO3 = ph_up_chalk_CaCO3)

# Resulting Water Profile:
# kable(chemistry)
mash_profile <- data.frame(Calcium=mash_Ca, Magnesium=mash_Mg, Sodium=mash_Na, Chloride=mash_Cl, Sulfate=mash_SO4)
mash_sparge_profile <- data.frame(Calcium=mash_sparge_Ca, Magnesium=mash_sparge_Mg, Sodium=mash_sparge_Na, Chloride=mash_sparge_Cl, Sulfate=mash_sparge_SO4)
palmers_ranges <- data.frame(Calcium="50 - 150", Magnesium="10 - 30", Sodium="0 - 150", Chloride="0 - 250", Sulfate="50 - 350")
```

<strong>Effective Alkalinity:</strong> `r effective_alkalinity`  
<strong>Residual Alkalinity:</strong> `r residual_alkalinity`  
<strong>Estimated Room-Temp Mash pH:</strong> `r estimated_room_temp_mash_ph`  
<strong>Desired Room-Temp Mash pH:</strong> 5.4 - 5.6  

Mash Profile:  
`r kable(mash_profile)`

Sparge Profile:  
`r kable(mash_sparge_profile)`

Palmer's Recommended Ranges:  
`r kable(palmers_ranges)`

## Yeast
```{r Yeast, echo= FALSE}

# Using style OG
# degree_plato <- (recipe$OG-1)*1000/4
# Using current predicted OG
degree_plato <- (predicted_og-1)*1000/4

liter_batch_size <- (batch_size+fermentation_loss)*3.785 # Liters in a gallon

cell_density <- (1e9*.75*degree_plato*liter_batch_size)

# Assumed that you want the right cell density and made it happen
pitched_cells <- cell_density

# IF otherwise, specify

# pitched_cells <- 150

yeast <- data.frame(
  recipe_id = 1,
  Yeast_Strain = "Fermentis Safale US-05",
  Attenuation = 81,
  Initial_Cell_Count_Billions = cell_density/1e9,
  Pitched_Cells_Billions = pitched_cells/1e9,
  Liters_For_Starter = 0,
  Time_For_Starter = 0
)


predicted_fg <- predicted_og-(yeast$Attenuation/100)*(predicted_og-1)
predicted_abv <- (1.05/0.79)*((predicted_og-predicted_fg)/predicted_fg)*100

kable(yeast[,2:7])
```

<strong>Predicted ABV:</strong> `r predicted_abv`

```{r Code to solve for starter volume needed, eval=FALSE, echo=FALSE}
# Code to solve for starter volume needed

I have some code that will bootleg this calculation and or some rules of thumb for starters...

```


## Fermentation

```{r Fermentation, echo= FALSE}
fermentation <- data.frame(
  recipe_id = 1,
  Temp1 = 62,
  Days1 = 28
)

kable(fermentation[,2:3])
```

Ferment at 62 ?F (17 ?C) until the beer attenuates fully. With healthy yeast, fermentation should be complete in a week, but don't rush it. The cooler than average ale fermentation temperature can extend the time it takes for complete attenuation. Rack to a keg and force carbonate or rack to a bottling bucket, add priming sugar, and bottle. Target a carbonation level of 2.5 to 3 volumes. [Fermentation source](https://byo.com/bock/item/1964-dunkelweizen-style-profile)

<!--* Note: [Hydration Instructions](http://howtobrew.com/book/section-1/yeast/preparing-yeast-and-yeast-starters)-->
* Note: [Hydration Instructions](http://www.fermentis.com/wp-content/uploads/2016/12/SafAle-US-053.pdf)  

## Brew Day Schedule

* Heat `r MV` Gallons up to `r IT` degrees Fahrenheit in Kettle pot.
* Add 5 g Gypsum, 4 g Calcium Chloride, 3 mL Lactic Acid to strike water.
* Slowly pour the grains into the water as to avoid clumping. Maintain +6 degrees F above mash temp (`r TMT` F) and then pour mash volume into mash tun. Place roasted pecans in grain bag and submerge the pecans in the mash. Check that the mash temp is `r TMT`. Hold here for 60 minutes. 
* Check mash temp after 20 minutes and if necessary, do a mini-decoction to bring temp back to `r TMT` F.
* While the grain mashes, measure `r MOV` gallons in lobster pot for the mash out volume. Begin heating to a boil.
* While the grain mashes, measure `r SVPMO` gallons in kettle pot for the rinse volume. Begin heating to 170 F.
* After mash completes, add mash out volume (`r MOV` gal) to mash tun. Hold at 170F for 10 minutes. 
* Now the mash should be at 170 degrees. We need to vorlauf. Be careful to pour gently and cycle the wort till it runs clean.  
* Now take the remaining `r SVPMO` gallons (rinse sparge) and slowly pour through the mash bed.
* Once fully drained, discard grains and set aside pecans (don't discard). 
* For the next 30-45 minutes bring to a boil. Once the wort is at 212 degrees F, set 30 minute timer and add the first hop addition (`r hops$Hop[1]`) at the beginning of the boil
* When timer goes off, set 15 minute timer and re-hydrate yeast:
    + Sprinkle the yeast in minimum 10 times its weight of sterile water or wort at 25 to 29°C (77°F to 84°F).
    + Total cell count = `r cell_density/1e9` billion (200 from pack 1, 139.94 from pack 2) 
    + Total cell weight = (139.94*11.5/200)+11.5 = 19.55 grams
    + 10 times its weight in water = 195.5 mL
* When timer goes off, set 5 minute timer and add the second hop addition (`r hops$Hop[2]`). 
* When timer goes off, set 10 minute timer, add Irish moss to boil, and tie pecan bag to side of kettle submerging them in boiling wort.
* Immediately gather sanitation and prepare buckets.
* When boil finishes, 
    + Remove pecans and transfer to a sanitized grain bag and place in fermentation bucket.
    + Gently stir yeast 
    + Cool wort till it reaches `r fermentation$Temp1` degrees F. Transfer to bucket and take OG reading.
* Pitch yeast and put beer in fridge for 14 days at `r fermentation$Temp1` degrees F.
* After 14 days, add purreed peaches and pecan vodka extract.
* After 7 days, taste and decide if using peach extract is necessary. Bottle. 
* After 7 days, drink.











