---
title: "Scottish 60/-"
author: '- BB; KS -'
date: "June 5th, 2016"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
```

```{r results="asis", echo = FALSE}
cat("
<style>
body {background-color: #FEF9E7;}
</style>
")
```

CONSTANTS

* Evaporation rate = 10%
* Boil time = 60 minutes
* Efficiency = 73.5%
* Boil kettle dead space = 0.125 gallons
* Lauter tun dead space = 0.125 gallons
* Fermentation loss = 0.125 gallons
* Target gravity = 1.0325 

MALT

* Golden Promise (Thomas Fawcett, 2.7째 L) - 5.8 lbs.
* Victory (Briess, 28째 L) - 0.23 lbs.
* Crystal (Great Western, 75째 L) - 0.07 lbs. 
* Roasted Barley (Thomas Fawcett, 440째 L) - 0.05 lbs.

HOPS

* Fuggle (UK) - 4.1% AA - 1 oz - 30 min boil (14 IBU) 
* Hallertau (German) - 2.5% AA - 0.5 oz - dry hop (0 IBU)
* East Kent Goldings (UK) - 4.5% AA - 0.5 oz - dry hop (0 IBU)

WATER

* Mash thickness = 1.25 qt/lbs
* Infusion water to 155F sacch. rest = 1.92 gallons @ 169 F
* No brewing salts or acids necessary
* Infusion water (mash-out) = 1.12 gallons @ 212 F
* Spare water = 4.75 gallons @ 180F

MASH SCHEDULE

* 155 F for 75 minutes (we meant to do only 60 minutes)
* 168 F mash-out

YEAST

* Yeast 1728, 1 pack = 100 billion cells
* Pitched directly into cooled wort @ 68 F

NOTES

* We ended up with EXACTLY five gallons of wort at 1.035 OG 
* Keep brewing constants as they are! We hit the holy grail on this one!

FERMENTATION

* Fermenting in basement in beginning of June. Hopefully it doesn't get too hot out there....
* Primary ferment one week
* Transfer, add dry hops, and secondary for two weeks
* Bottle for one week